---
title: "Sentiment Analysis"
author: "Sarasati Palawita"
date: "16 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) #Data manipulation (also included in the tidyverse package)
library(tidytext) #Text mining
library(tidyr) #Spread, separate, unite, text mining (also included in the tidyverse package)
library(RTextTools)
library(readxl) #To read excel
library(stringr)
library(e1071) #To run SVM
library(caret) #Control

#Visualizations!
library(ggplot2) #Visualizations (also included in the tidyverse package)
library(ggrepel) #`geom_label_repel`
library(gridExtra) #`grid.arrange()` for multi-graphs
library(knitr) #Create nicely formatted output tables
library(kableExtra) #Create nicely formatted output tables
library(formattable) #For the color_tile function
```

Define color
```{r }
#Define some colors to use throughout
my_colors <- c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#FF1234", "#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#FF1234")

#Customize ggplot2's default theme settings
theme_lyrics <- function(aticks = element_blank(),
                         pgminor = element_blank(),
                         lt = element_blank(),
                         lp = "none")
{
  theme(plot.title = element_text(hjust = 0.5), #Center the title
        axis.ticks = aticks, #Set axis ticks to on or off
        panel.grid.minor = pgminor, #Turn the minor grid lines on or off
        legend.title = lt, #Turn the legend title on or off
        legend.position = lp) #Turn the legend on or off
}

#Customize the text tables for consistency using HTML formatting
my_kable_styling <- function(dat, caption) {
  kable(dat, "html", escape = FALSE, caption = caption) %>%
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"),
                full_width = FALSE)
}

```

Data Retrieve
```{r pressure, echo=FALSE}
setwd("C:/Users/Sarasati Palawita/Documents/Personal/BIPM 2017/Master Thesis 2018/Data")
cleaned_data = read_excel("cleaned_data_no_duplicate.xlsx")
glimpse(charlie) #Transposed version of `print()`

```

Data Partition
```{r}
cleaned_data$lyrics <- as.character(cleaned_data$lyrics)
cleaned_data = na.omit(cleaned_data)
cleaned_data <- as_tibble(cleaned_data)
class(cleaned_data)

data <- cleaned_data

train_source <- cleaned_data %>%
  group_by(genre) %>% 
  nest() %>%            
  mutate(n = c(1250)) %>% 
  mutate(samp = map2(data, n, sample_n)) %>% 
  select(genre, samp) %>%
  unnest()
```

Data Conditioning
```{r lyrics_song$lyrics}
# function to expand contractions in an English-language source
fix.contractions <- function(doc) {
  # "won't" is a special case as it does not expand to "wo not"
  doc <- gsub("won't", "will not", doc)
  doc <- gsub("can't", "can not", doc)
  doc <- gsub("n't", " not", doc)
  doc <- gsub("'ll", " will", doc)
  doc <- gsub("'re", " are", doc)
  doc <- gsub("'ve", " have", doc)
  doc <- gsub("'m", " am", doc)
  doc <- gsub("'d", " would", doc)
  # 's could be 'is' or could be possessive: it has no expansion
  doc <- gsub("'s", "", doc)
  return(doc)
}
train_source$lyrics <- sapply(train_source$lyrics, fix.contractions)
```

To get rid of special characters
```{r removeSpecialChars}
# function to remove special characters
removeSpecialChars <- function(x) gsub("[^a-zA-Z0-9 ]", " ", x)
# remove special characters
train_source$lyrics <- sapply(train_source$lyrics, removeSpecialChars)
# convert everything to lower case
train_source$lyrics <- sapply(train_source$lyrics, tolower)
```

Undesirable words
```{r undesirable_words}
undesirable_words <- c("chorus", "lyrics", "intro", "clichÃfÂ©s",
                       "theres", "bridge", "fe0f",
                       "chorus", "verse", "[chorus]", "[verse]",
                       "2", "2x", "3x", "1x", 
                       "4", "ooh", "uurh", "uuh", "pheromone", "poompoom", "3121", 
                       "matic", " ai ", " ca ", " la ", " na ", 
                       " da ", " uh ", " tin ", "  ll", "transcription", "YoncÃfÂ©", "BeyoncÃfÂ©")

```

Tokenization
```{r}
#Create tidy text format: Unnested, Unsummarized, -Undesirables, Stop and Short words
train_tidy <- train_source %>%
  unnest_tokens(word, lyrics) %>% #Break the lyrics into individual words
  filter(!word %in% undesirable_words) %>% #Remove undesirables
  filter(!nchar(word) < 3) %>% #Words like "ah" or "oo" used in music
  anti_join(stop_words) #Data provided by the tidytext package
```

Create Sentiment Dataset
```{r}
train_bing <- train_tidy %>%
  inner_join(get_sentiments("bing"))

train_nrc <- train_tidy %>%
  inner_join(get_sentiments("nrc"))

train_nrc_sub <- train_tidy %>%
  inner_join(get_sentiments("nrc")) %>%
  filter(!sentiment %in% c("positive", "negative"))
```

1. Genre Analysis with NRC Sentiment
```{r}
plot_pop <- train_nrc %>%
  filter(genre == "Pop") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle(train_nrc$genre) +
  coord_flip()

plot_jazz <- train_nrc %>%
  filter(genre == "Jazz") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Jazz") +
  coord_flip()

plot_rock <- train_nrc %>%
  filter(genre == "Rock") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Rock") +
  coord_flip()

plot_indie <- train_nrc %>%
  filter(genre == "Indie") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Indie") +
  coord_flip()

plot_metal <- train_nrc %>%
  filter(genre == "Metal") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Metal") +
  coord_flip()

plot_hiphop <- train_nrc %>%
  filter(genre == "Hip-Hop") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Hip-Hop") +
  coord_flip()

plot_rb <- train_nrc %>%
  filter(genre == "R&B") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("R&B") +
  coord_flip()

plot_electro <- train_nrc %>%
  filter(genre == "Electronic") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Electronic") +
  coord_flip()

plot_country <- train_nrc %>%
  filter(genre == "Country") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Country") +
  coord_flip()

plot_folk <- train_nrc %>%
  filter(genre == "Folk") %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle("Folk") +
  coord_flip()

grid.arrange(plot_pop, plot_jazz, ncol = 2)
grid.arrange(plot_metal, plot_rock, ncol = 2)
grid.arrange(plot_hiphop, plot_rb, ncol = 2)
grid.arrange(plot_country, plot_folk, ncol = 2)
grid.arrange(plot_indie, plot_electro, ncol = 2)
grid.arrange(plot_pop, plot_jazz, plot_metal, plot_rock, plot_hiphop, plot_rb, plot_country, plot_folk, plot_indie, plot_electro, ncol = 2)
```
Relationship between Mood and Genre
```{r}
grid.col = c("1970s" = my_colors[1], "1980s" = my_colors[2], "1990s" = my_colors[3], "2000s" = my_colors[4], "2010s" = my_colors[5], "anger" = "grey", "anticipation" = "grey", "disgust" = "grey", "fear" = "grey", "joy" = "grey", "sadness" = "grey", "surprise" = "grey", "trust" = "grey")

genre_mood <-  train_nrc %>%
  filter(genre != "NA" & !sentiment %in% c("positive", "negative")) %>%
  count(sentiment, genre) %>%
  group_by(genre, sentiment) %>%
  summarise(sentiment_sum = sum(n)) %>%
  ungroup()

circos.clear()
#Set the gap size
circos.par(gap.after = c(rep(5, length(unique(genre_mood[[1]])) - 1), 15,
                         rep(5, length(unique(genre_mood[[2]])) - 1), 15))
chordDiagram(genre_mood, grid.col = grid.col, transparency = .2)
title("Relationship Between Mood and Genre")
```

2.	Genre Polarity Analysis with Bing Sentiment
```{r}
#View(train_bing)
train_polarity_genre <- train_bing %>%
  count(sentiment, genre) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(polarity = positive - negative,
    percent_positive = positive / (positive + negative) * 100)

train_polarity_genre <- train_polarity_genre %>%
  mutate(genre = ifelse(genre == "Pop", "Pop",
                 ifelse(genre == "Jazz", "Jz",
                 ifelse(genre == "Rock", "Rck",
                 ifelse(genre == "Hip-Hop", "Hp",
                 ifelse(genre == "Metal", "Met",
                 ifelse(genre == "Electronic", "El",
                 ifelse(genre == "R&B", "R&b",
                 ifelse(genre == "Indie", "In",
                 ifelse(genre == "Country", "Cty",
                 ifelse(genre == "Folk", "Flk", genre)))))))))))
  
#Polarity by chart
plot1 <- train_polarity_genre %>%
  ggplot( aes(genre, polarity, fill = genre)) +
  geom_col() +
  scale_fill_manual(values = my_colors[1:10]) +
  geom_hline(yintercept = 0, color = "red") +
  theme_lyrics() + theme(plot.title = element_text(size = 8)) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Polarity By Genre")

#Percent positive by chart
plot2 <- train_polarity_genre %>%
  ggplot( aes(genre, percent_positive, fill = genre)) +
  geom_col() +
  scale_fill_manual(values = c(my_colors[1:10])) +
  geom_hline(yintercept = 0, color = "red") +
  theme_lyrics() + theme(plot.title = element_text(size = 7)) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Percent Positive By Genre")
grid.arrange(plot1, plot2, ncol = 2)

```

3.	Human Interpretation vs Sentiment Analysis with NRC Sentiment
Charlie - Part 1 - Nrc- We don't talk anymore
```{r}
#1. Retrieve Data
setwd("C:/Users/Sarasati Palawita/Documents/Personal/BIPM 2017/Master Thesis 2018/Data")
charlie = read_excel("Charlie_Puth.xlsx")
#Convert to characted type
charlie$lyric <- as.character(charlie$lyric)
#Convert to Tibble data
charlie <- as_tibble(charlie)
class(charlie)

#2. Function to remove special characters
  removeSpecialChars <- function(x) gsub("[^a-zA-Z0-9 ]", " ", x)
  # remove special characters
  charlie$lyrics <- sapply(charlie$lyrics, removeSpecialChars)
  # convert everything to lower case
  charlie$lyrics <- sapply(charlie$lyrics, tolower)

#3. Tokenization
charlie_tidy <- charlie %>%
  unnest_tokens(word, lyrics) %>% #Break the lyrics into individual words
  filter(!word %in% undesirable_words) %>% #Remove undesirables
  filter(!nchar(word) < 3) %>% #Words like "ah" or "oo" used in music
  anti_join(stop_words) #Data provided by the tidytext package

#4. Get NRC sentiments
charlie_nrc <- charlie_tidy %>%
  inner_join(get_sentiments("nrc"))

#5. Visualize the NRC Graph
charlie_nrc %>%
  group_by(sentiment) %>%
  summarise(word_count = n()) %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_count, fill = -word_count)) +
  geom_col() +
  guides(fill = FALSE) +
  theme_minimal() + theme_lyrics() +
  labs(x = NULL, y = "Word Count") +
  ggtitle(charlie_tidy$song) +
  coord_flip()
```

Charlie - Part II - NRC Detail Information
```{r}
#6. Visualize the NRC Detail Table
charlie_tidy %>%
  distinct(word) %>%
  inner_join(get_sentiments("nrc")) %>%
  ggplot(aes(x = word, fill = sentiment)) +
  facet_grid(~sentiment) +
  geom_bar() + #Create a bar for each word per sentiment
  theme_lyrics() +
  theme(panel.grid.major.x = element_blank(),
        axis.text.x = element_blank()) + #Place the words on the y-axis
  xlab(NULL) + ylab(NULL) +
  ggtitle(charlie_tidy$song) +
  coord_flip()
```